### 2021-12-08 documents/sections/extensions.tex
yukpa trimming

\subsubsection{\yukpa \obj{j-}}
\label{sec:yukpa}
The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
However, it does preserve the \setone prefixes in the immediate past, where a unified intransitive paradigm \pcref{tab:yukreg} is found.
The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{s_a_} \obj{m(ɨ)-} to (now former) \gl{s_p_} verbs like \obj{nɨ} \qu{to sleep}.
%
\input{floats/yukreg}%
%
These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their semantics and the reflex of \detrz.
The same prefix occurs on transitive verbs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, meaning that \obj{j(ɨ)-} is the reflex of the \PC \gl{1}\gl{s_p_} marker \rc{u(j)-}.
On the other hand, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.

\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
\a<yuk-5>
\begingl
\gla aw j-esare//
\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
\glft \qu{S/he saw me.}//
\endgl
\a<yuk-1>
\begingl
\gla aw {\normalfont ∅}-esare//
\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
\glft \qu{I saw it.}//
\endgl
\xe
%
Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and since it is prone to phonological erosion elsewhere in the family \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identifiedn  as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.

In intransitive verbs, this first-person zero marking is attested in a single verb, \obj{to} \qu{to go} \exref{yuk-7}.
It diverges from regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} in \cref{tab:yukreg}.
It is thus the only verb attested as having resisted the extension of \obj{j-} in \yukpa.

>>>>

\subsubsection{\yukpa \obj{j-}}
\label{sec:yukpa}
The divergent nature of the family-internal isolate \yukpa is \textit{inter alia} visible in the loss of many \setone forms and the formation of non-cognate innovative constructions \parencite{meira2006syntactic}.
However, it does preserve the \setone prefixes in the immediate past, where a unified intransitive paradigm \pcref{tab:yukreg} is found.
The wholesale loss of \gl{1+2} as an inflectional value was combined with the extension of \gl{2}\gl{s_a_} \obj{m(ɨ)-} to (now former) \gl{s_p_} verbs like \obj{nɨ} \qu{to sleep}.
%
\input{floats/yukreg}%
%
These verbs share their first person marker \obj{j(ɨ)-} with former \gl{s_a_} verbs like \obj{otum} \qu{to wash self}, identifiable by their semantics and the reflex of \detrz.
The same prefix occurs on transitive verbs in \gl{3}>\gl{1} scenarios \exref{yukpatr.yuk-5}, so \obj{j(ɨ)-} is the reflex of the \PC \gl{1}S\textsubscript{(P)} marker \rc{u(j)-}.
In contrast, \gl{1}>\gl{3} scenarios are zero-marked \exref{yukpatr.yuk-1}.

\pex<yukpatr>\yukpa \parencite[][139]{meira2006syntactic}
\a<yuk-5>
\begingl
\gla aw j-esare//
\glb \gl{1}\gl{pro} \gl{3}>\gl{1}-see//
\glft \qu{S/he saw me.}//
\endgl
\a<yuk-1>
\begingl
\gla aw {\normalfont ∅}-esare//
\glb \gl{1}\gl{pro} \gl{1}>\gl{3}-see//
\glft \qu{I saw it.}//
\endgl
\xe
%
Since \PC \gl{1}\gl{s_a_} \rc{w(ɨ)-} was extended to \gl{1}>\gl{3} scenarios in most languages \parencite[81--82]{gildea1998}, and given its inclination for phonological erosion elsewhere in the family \pcref{sec:pekodian,sec:waiwaian}, the zero marking in \gl{1}>\gl{3} scenarios can be identified as the \yukpa reflex of \rc{w-}, contrasting with \obj{j(ɨ)-}, the reflex of \rc{u(j)-}.

In intransitive verbs, this first-person zero marking is attested in a single verb, \obj{to} \qu{to go} \exref{yuk-7}.
It diverges from regular C-initial verbs, which take \obj{jɨ-}, like \qu{to sleep} \pcref{tab:yukreg}.
It can thus be identified as having resisted the extension of \obj{j-} in \yukpa.
